---
title: "Assignment3"
output: pdf_document
date: "2024-11-13"
---
Step 1: Stimulating some data
1. set arbitrary values:
drift rate: 0
boundary separation: 1.5
nondecision time: 0.3
2. these values will be the population mean for each EZ-DDM parameter
```{r}
drift_mean<- 0
bound_mean<- 1.5
nondt_mean<- 0.3
```
3. build a function that generates the individual-level parameters 
```{r}
generate_individual_parameters<-function(n_Participants, drift_mean, bound_mean, nondt_mean){
  drift_sd<-1
  bound_sd<-0.5
  nondt_sd<-0.1
  drift_rate<- rnorm(n_Participants, mean=drift_mean, sd=drift_sd)
  bound_values<- rnorm(n_Participants, mean=bound_mean, sd=bound_sd)
  nondt_values<- rnorm(n_Participants, mean=nondt_mean, sd=nondt_sd)
data.frame(Participant = 1:n_Participants, drift_rate, bound_values, nondt_values)
}

```
4. Build a function that generates summary stats 
```{r}
summary_stats<- function(individual_parameter, n_trials){
  
  a<-individual_parameter$bound_values
  v<-individual_parameter$drift_rate
  t<-individual_parameter$nondt_values
  
  #equation 1
  A_pred_mean<- 1 / (1 + exp(-a * v))
  #equation 2
  RT_pred_mean<-t + ((a / (2 * v^3 )) * ((exp(-a * v) - 1) / (exp(-a * v) + 1)))
  #equation 3
  RT_pred_var <- (a / (2 * v^3)) * (-2 * a * v * exp(-a * v) - exp(-2 * a * v) + 1) / (exp(-a * v) + 1)^2

  #equation 4
  A_obs_total <- rbinom(length(A_pred_mean), size = n_trials, prob = A_pred_mean)
  #equation 5
  RT_obs_mean <- rnorm(length(RT_pred_mean), mean=RT_pred_mean, sd=sqrt((RT_pred_var/n_trials)))
  #equation 6
  RT_obs_var <- rnorm(length(RT_pred_var), mean=RT_pred_var, sd = sqrt((2*RT_pred_var^2)/(n_trials-1)))
  
 data.frame(Participant = individual_parameter$Participant, A_pred_mean, RT_pred_mean, RT_pred_var, A_obs_total, RT_obs_mean, RT_obs_var)
  }
```
5. test using 50 participants and 150 trials each
```{r}
set.seed(202)
n_Participants<-80
n_trials<-200

indiv_parameters<-generate_individual_parameters(n_Participants, drift_mean, bound_mean, nondt_mean)

summary_table<- summary_stats(indiv_parameters, n_trials)


```


Step 2:
JAGS module 
```{r}

write("
model {
  bound_mean ~ dnorm(1.5, pow(0.05, -2))T(0.1, 5.0)  
  nondt_mean ~ dnorm(0.3, pow(0.02, -2))T(0.1, 0.6) 
  drift_mean ~ dnorm(0, pow(0.5, -2))T(-5.0, 5.0)    
  bound_sdev ~ dunif(0.1, 1)    
  nondt_sdev ~ dunif(0.05, 0.5)   
  drift_sdev ~ dunif(0.1, 1.5) 

  # Sampling model
  for (p in 1:n_Participants) {
    bound[p] ~ dnorm(bound_mean, pow(bound_sdev, -2))T(0.10, 5.00)
    nondt[p] ~ dnorm(nondt_mean, pow(nondt_sdev, -2))T(0.05,)
    drift[p] ~ dnorm(drift_mean, pow(drift_sdev, -2))T(-5.00, 5.00)

    # Forward equations from EZ Diffusion
    ey[p] = exp(-bound[p] * drift[p])
    Pc[p] = 1 / (1 + ey[p])
    PRT[p] = 2*pow(drift[p],3)/bound[p]*pow(ey[p] + 1, 2)/(2 *-bound[p]*              drift[p] * ey[p] - ey[p]*ey[p]+1)
    MDT[p] = (bound[p] / (2 * drift[p])) * (1 - ey[p]) / (1 + ey[p])
    MRT[p] = MDT[p] + nondt[p]

    # Loss functions using MRT, PRT, and Pc
    correct[p] ~ dbin(Pc[p], n_trials)
    meanRT[p] ~ dnorm(MRT[p], PRT[p] * n_trials)
    varRT[p] ~ dnorm(1 / PRT[p], 0.5 * (n_trials - 1) * PRT[p] * PRT[p])
  }
}
", "./model.bug")


```

Step 3: JAGS set-up
```{r}
parameters<-c("drift_mean", "bound_mean", "nondt_mean")
n.iter<-1500
n.chains<-4
n.burnin<-500
n.thin<-1

data_toJAGS<- list(correct=summary_table$A_obs_total,
                   meanRT=summary_table$RT_obs_mean,
                   varRT=summary_table$RT_obs_var,
                   n_Participants=n_Participants,
                   n_trials=n_trials 
)

myinits<-rep(list(list()), n.chains)
set.seed(202)
for(i in 1:n.chains){
  myinits[[i]]<- list(drift=rnorm(n_Participants, 0, 0.5))
  
}

```
Step 4: Run JAGS
```{r}
library(rjags)
library(coda)
library(R2jags)
samples <- jags(data = data_toJAGS, parameters.to.save = parameters,
                model= "./model.bug",
                n.chains = n.chains, n.iter = 1000,
                n.burnin = 100, n.thin = n.thin,DIC = T, inits = myinits)
```

step 6: graph

```{r}

library(ggplot2)
drift_mean_samples <- as.vector(samples$BUGSoutput$sims.array[,, "drift_mean"])
bound_mean_samples <- as.vector(samples$BUGSoutput$sims.array[,, "bound_mean"])
nondt_mean_samples <- as.vector(samples$BUGSoutput$sims.array[,, "nondt_mean"])

posterior_data <- data.frame(
  drift_mean = drift_mean_samples,
  bound_mean = bound_mean_samples,
  nondt_mean = nondt_mean_samples
)

ggplot(posterior_data, aes(x = drift_mean))+
  geom_density(fill = "skyblue", alpha = 0.6) +
  labs(title = "Posterior Distribution of Drift Mean", x = "Drift Mean", y = "Density") + geom_vline(xintercept =0 , color = "red")
ggplot(posterior_data, aes(x = bound_mean))+
  geom_density(fill = "pink", alpha = 0.6) +
  labs(title = "Posterior Distribution of Bound Mean", x = "Bound Mean", y = "Density") + geom_vline(xintercept =1.5 , color = "red")
ggplot(posterior_data, aes(x = nondt_mean))+
  geom_density(fill = "purple", alpha = 0.6) +
  labs(title = "Posterior Distribution of NonDt Mean", x = "NonDt Mean", y = "Density") + geom_vline(xintercept =0.3 , color = "red")

```



